---
format: 
  html: 
    self-contained: true
    toc: true             
  pdf: 
    number-sections: true 
    toc: true            
    pdf-engine: xelatex 
editor: visual
---

## Bivariate Data Visualizations

```{r}
library(ggplot2)
library(dplyr)
library(papaja)
```

```{r}
#| echo: false
#| message: false
#| warning: false
source("loadFilteredData.R")
```

### Figure 1 Scatterplot

```{r}
#| message: false
#| warning: false
ggplot(data=df) +
  geom_point(aes(x=IUS_Total, 
                 y=STARS_Total)) +
  geom_smooth(aes(x=IUS_Total, 
                  y=STARS_Total), 
                  method="lm") +
  ggtitle("IUS x STARS") +
  xlab("IUS") +
  ylab("STARS") +
  theme_apa()
```

My independent variable is intolerance for uncertainty (IUS_Total), and my dependent variable is statistics anxiety (STARS_Total).

In the figure above, there is a positive relationship between a statistics anxiety and intolerance for uncertainty. As intolerance for uncertainty increases, statistics anxiety tends to also increase.

### Figure 2 Scatterplot

```{r}
#| message: false
#| warning: false
ggplot(data=df) +
  geom_point(aes(x=IUS_Total, 
                 y=STARS_Total,
                 color=ATMS_Persistence_Total)) +
  geom_smooth(aes(x=IUS_Total, 
                  y=STARS_Total), 
                  method="lm") +
  ggtitle("IUS x STARS") +
  xlab("IUS") +
  ylab("STARS") +
  theme_apa()
```

I am adding persistence (ATMS_Persistence_Total) as my third variable. The colors seem widely spread throughout the graph with no obvious pattern.

### Extra example with facet wrap

```{r}
# univariate graph (categorical variable)
df %>% 
  filter(!is.na(gender)) %>% 
  ggplot(aes(x=age))+
  geom_bar()+
  ggtitle("Age of Participants") +
  facet_wrap(~gender)
```
