---
title: "intro"
format: 
  html: 
    self-contained: true
  pdf: 
    number-sections: true 
    toc: true            
    pdf-engine: xelatex 
editor: visual
---

# Introduction

## Comments

```{r}
# This is a comment
```

## R as a calculator

Simple:

```{r}
10 + 20
```

Order of operations, two examples:

```{r}
1 + 2 * 4
(1 + 2) * 4
```

## Variables

For beginners: use only `<-`. Drop `->` entirely.

```{r}
sales <- 350
royalty <- 7
revenue <- sales * royalty
revenue
```

Updating a variable is good and should stay:

```{r}
revenue <- revenue + 550
revenue
```

If you are using the console, you can list objects with `objects()` (the document also highlights using RStudio’s Environment panel as an easy option).

```{r}
objects()
```

The document’s core warning: once you remove variables, there’s no undo unless you saved them.

```{r}
rm(revenue, royalty)
```

### Functions

Remove extra spacing inside parentheses. Beginners copy patterns.

```{r}
sqrt(225) 
abs(-13) 
round(3.14159, 2)
```

Named arguments are useful later, but early on it is more typing than learning. If you want to mention them, show one example and move on:

```{r}
round(x = 3.14159, digits = 2)
```

### Characters

```{r}
greeting <- "hello"
greeting
greeting[1]
nchar(greeting)
```

### Logical Values

Strongly recommend you avoid teaching `T` and `F`. They are changeable. Stick to `TRUE` and `FALSE`.

```{r}
2 + 2 == 4
2 + 2 == 5
```

### NaN (Not a number)

A common way to produce `NaN` is 0 divided by 0.

```{r}
0 / 0
```

### NA (missing)

`NA` indicates missing information (very common with real datasets). The document contrasts NA with NaN: NaN is an undefined math result, while NA is missing information.

```{r}
Q <- c(1, NA, 3)
Q
```

### NULL (no value)

`NULL` means “no value at all,” which the document treats as a stronger absence than NA.

```{r}
z <- NULL
z
```

### Vectors

One type only (all numeric, or all characters, or all logical, etc.)

```{r}
months <- month.name
months
```

```{r}
v1 <- c(1, 2, 3)              # numeric vector
v2 <- c("a", "b", "c")        # character vector
v3 <- c(TRUE, FALSE, TRUE)    # logical vector
```

### List

Can mix types.

```{r}
example_list <- list(
  name = "Ana",
  scores = c(80, 90, 85),
  passed = TRUE
)
```

### Factors

Numeric codes for nominal groups can lead to nonsense math (like inventing groups that do not exist), so convert to a factor.

```{r}
group <- c(1,1,1,2,2,2,3,3,3)
group + 2   # misleading outcome for nominal data (demonstration)
```

Convert to factor:

```{r}
group <- as.factor(group)
class(group)
```

This is the exact workflow shown.

Then try the “bad math” again:

```{r}
group + 2
```

### Label factor levels

Assign meaningful labels with `levels(...) <- ...`.

```{r}
levels(group) <- c("group 1", "group 2", "group 3")
group

```

Example for gender coding:

```{r}
gender <- as.factor(c(1,1,1,1,1,2,2,2,2))
levels(gender) <- c("male", "female")
gender
```

Acknowledgement: [learningstatisticswithr](https://learningstatisticswithr.com/book/introR.html)
